#Lab_04 {lab_04}

rob wells, 8/14/2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)
# Load janitor here
library(janitor)


```

### Load data

In this data set, each row represents a city or town in Maryland with the 2010 and 2020 figures for: 

-  Total population 
-  White only population
-  Black or African American only population
-  Hispanic only population 
-  Native American only population
-  Asian only population
-  Pacific Islander (including native Hawiian) only population

Background on this data from [Social Explorer](https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001) 
```{r}
# Write code to load the Maryland race by city, place for 2010 - 2020 and call it md_race.
# The data is called lab_03.csv, and is in the same folder as lab_03.qmd (this data notebook)
md_race<- read_csv("lab_04.csv") %>% 
  as.data.frame()
```

## Answer questions

Q1: Name the five cities that had the largest Hispanic population as a percentage of the entire population in 2020 and call it top5_hispanic. Produce a top 5 table for each, selecting the place, 2020 Hispanic, 2020 total and your new calculation.

A1:
```{r}

# Write your code in this codeblock.
md_race <- md_race %>% 
  mutate(percentof_total = (x2020_hispanic/x2020_total))
top5_hispanic <- md_race %>% 
  select(place, x2020_hispanic, x2020_total, percentof_total) %>% 
  slice_max(percentof_total, n=5)


```

Q2: Compare the top 10 cities with the largest Black population as a percentage of the overall population in 2010 and 2020. Filter for places with more than 20,000 population in 2010. Call your results pct_black. Compare results for 2020 vs 2010. What is different? What surprised you? 

A2:Half of the top 10 on the list actually fell in their percent of Black residents compared to the whole population. It was surprising given that they still are some of the highest Black populated areas, yet their percent of the total population is still down from a decade ago. 

```{r}
#write your code here
md_race <- md_race %>% 
  mutate(x2020blackpctof_total = (x2020_black/x2020_total)) %>% 
  mutate(x2010blackpctof_total = (x2010_black/x2010_total))
pct_black <- md_race %>% 
  select(place, x2020_black, x2010_black, x2020_total, x2010_total, x2020blackpctof_total, x2010blackpctof_total) %>% 
  filter(x2010_total >20000) %>% 
  slice_max(x2020blackpctof_total, n=10)

```

Q3: Find the 15 communities that had the great gains in Hispanic population from 2010 to 2020 as measured by percentage change. Also calculate percentage change for the community's overall growth during the same period. Filter your results for communities with more than 10,000 people as of 2020. Your results will be hisp_growth. Look at your results, craft a one-paragraph story pitch for your editor. 

A3: Hispanic population is growing in some communities that are declining overall. Overia CDP, for example, grew their Hispanic population by more than 400% over the last decade, despite their overall population declining by 4%.Even in places that aren't declining in population, there is a much sharper rise in the Hispanic numbers than there are in the percent increase in total population, meaning more Hispanics are coming in while others are moving out. 

```{r}
md_race <- md_race %>% 
  mutate(hispanicpct_change = (x2020_hispanic-x2010_hispanic)/x2010_hispanic) %>% 
  mutate(totalpct_change = (x2020_total-x2010_total)/x2010_total)
hisp_growth <- md_race %>% 
  select(place, x2020_hispanic, x2010_hispanic, x2020_total, x2010_total, hispanicpct_change, totalpct_change) %>% 
  filter(x2020_total >10000) %>% 
  slice_max(hispanicpct_change, n=15)


```
