#Lab_03 {lab_03}

rob wells, 8/14/2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)
# Load janitor here
library(janitor)

```

### Load data

In this data set, each row represents a city or town in Maryland with the 2010 and 2020 figures for: 

-  Total population 
-  White only population
-  Black or African American only population
-  Hispanic only population 
-  Native American only population
-  Asian only population
-  Pacific Islander (including native Hawiian) only population

Background on this data from [Social Explorer](https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001) 
```{r}
# Write code to load the Maryland race by city, place for 2010 - 2020 and call it md_race.
# The data is called lab_03.csv, and is in the same folder as lab_03.qmd (this data notebook)

# your answer here
md_race <- rio::import("https://docs.google.com/spreadsheets/d/1F6V_LH9_NWl2Ed60PA2wRZoHgNxRuNJs7C1NGesmo_I/edit#gid=260940342", which = "MDRace")
```


## Answer questions

Q1: Which five cities have the largest overall Hispanic population in 2020? Name the five cities that had the largest Hispanic population in 2020. 

A1:
```{r}

# Write your code in this codeblock.
md_race %>% 
  select(place, x2020_hispanic) %>% 
  slice_max(x2020_hispanic, n = 5)
# your answer here
#1.Baltimore city	32627		
#2.Wheaton CDP	22787		
#3.Germantown CDP	21381		
#4.Silver Spring CDP	21087		
#5.Aspen Hill CDP	18635		
```

Q2: Find the cities with the largest Black population and arrange the results in descending order. 

A2:

```{r}
#write your code here
md_race %>% 
  select(place, x2020_black) %>%
  slice_max(x2020_black, n = 5)%>%
  arrange(desc(x2020_black)) 
# your answer here
#1.Baltimore city	375002			
#2.Waldorf CDP	47903			
#3.Bowie city	32784			
#4.Clinton CDP	30679			
#5.Columbia CDP	28991	
```

Q3: Create a table called top_hispanic that shows the top 15 communities with the largest Hispanic population in 2020, and also display the 2010 Hispanic population and the total community population for 2020. Look at your results - what stands out to you? Craft a one-paragraph story pitch for your editor. 

A3: The Hispanic population in the state of Maryland is on the rise. All of the top 15 places in Maryland with the most Hispanic people have grown over the last decade. However, they remain the minority in most communities. We can write about the rise of Hispanic living in the state, and how they are handling being in a place of minority despite their rise in population. 

```{r}

# your answer here
md_race %>% 
  select(place, x2020_hispanic, x2010_hispanic, x2020_total) %>% 
 group_by(place) %>% 
 slice_max(x2020_hispanic, n = 15)%>%
  arrange(desc(x2020_hispanic))
```
