#lab_08
Sean Mussenden, Aug 16, 2022

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load janitor and the tidyverse
```{r}
library(tidyverse)
library(janitor)
```

## Load data
```{r}
maryland_population <- read_csv("data/maryland_population.csv")
maryland_hispanic <- read_csv("data/maryland_hispanic.csv")
maryland_black <- read_csv("data/maryland_black.csv")
maryland_white <- read_csv("data/maryland_white.csv")
```

**Task**: Load four tables of U.S. Census data.  Each table has one row per Maryland county with 2020 population information.  Each table has these same two columns:

* geoid - a unique identifier for each county
* name - the county name

The third column in each table represents the total population in one of four demographic categories.

* maryland_population.csv - total_pop - total population 
* maryland_hispanic.csv - hispanic_pop - total Hispanic population
* maryland_white.csv - white_pop - total White alone population
* maryland_black.csv - black_pop - total Black alone population

Some of the information in the tables may need to be cleaned or validated in order to produce the correct answer.  

## Questions

**Q1**  What Maryland county has the largest Hispanic population (by percentage of total population)? 
**A1** 
```{r}
maryland_hispanic <- maryland_hispanic %>%
  mutate(name = case_when(
    name == "Montgomery County,Maryland" ~ "Montgomery County, Maryland",
    TRUE ~ name
  ))
```

```{r}

  maryland_hispanic_total <- maryland_population %>%
  inner_join(maryland_hispanic, by=c("geoid", "name"))
```

```{r}
maryland_hispanic_total %>%
  select(geoid, name, total_pop, hispanic_pop) %>%
    mutate(pct_of_total = (hispanic_pop/total_pop)*100) %>% 
    arrange(desc(pct_of_total))
```
A: Montgomery County has the highest Hispanic population by total, with Hispanics accounting for 19.5% of the total population of the county.

**Q2**  What is the total population of people who identify as white alone in Maryland? 
**A2**  
```{r}
maryland_white <- maryland_white %>%
  mutate(white_pop = case_when(
    name == "Somerset County, Maryland" ~ 13510,
    TRUE ~ white_pop
  ))
```

```{r}
maryland_white %>% 
summarise(sum(white_pop))
```
A: There are 3,275,048 people that identify as White in Maryland. 
**Q3** What is the average Black population in all Maryland counties? 


**A3** 

```{r}
maryland_black %>%
  summarise(
    average_black_population = mean(black_pop, na.rm=TRUE)
  )
```

The average Black population in Maryland is 53864.91. 